Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-18 latex/main.tex
Updates to copying patterns and QC

Before characterizing recombination events in detail,
we sought to remove probable artifactual recombinants
due to correlated bases, which can be caused by
systematic errors in genome sequences
associated with sequencing protocols \cite{Turakhia2020},
multi-nucleotide substitutions (MNS) \cite{DeMaio2024}, and
alignment errors (e.g., deletions due to MAFFT misalignment \cite{Hunt2024}).

We characterized artifactual recombination events
as having fewer than four such loci supporting the left or right parent
on either side of a suggested breakpoint.
We treated adjacent and near-adjacent sites
(separated by at most three bases) as single loci.
Also, when counting loci supporting a focal parent,
we penalized loci where the character(s)
in the recombinant matches that in the non-focal parent
but not in the focal parent.
Illustrative copying patterns in Figure~\ref{fig:copying_pattern_examples}
show the evidence for some QC-failing recombinants:
(1) Q1, 686;
(2) Q3, MNS at 13423-13424 \cite{DeMaio2024};
(3) Q4, 22194 (possibly MAFFT alignment error \cite{Hunt2024}).

Of the 929 recombination events in the ARG,
this measure identifies 543 as artifactual,
leaving 386 ``robust'' events (Figure~\ref{fig:recombinant_qc}).
Evidence that this filter effectively flagged artifacts can be seen
in the clear association between the AmpliSeq V1 kit and
recombination events that we identified as artifactual.

>>>>

Before characterizing recombination events in detail,
we sought to remove probable artifactual recombinants
due to correlated bases, which can be caused by
systematic errors in genome sequences
associated with sequencing protocols \cite{Turakhia2020},
multi-nucleotide substitutions (MNS) \cite{DeMaio2024}, and
alignment errors \cite{Hunt2024}).

We characterized artifactual recombination events
as having fewer than four such loci supporting the left or right parent
on either side of a suggested breakpoint.
We treated adjacent and near-adjacent sites
(separated by at most three bases) as single loci.
Also, when counting loci supporting a focal parent,
we penalized loci where the character(s)
in the recombinant matches that in the non-focal parent
but not in the focal parent.

Of the 929 recombination events in the ARG,
this measure identifies 543 as artifactual,
leaving 386 ``robust'' events (Figure~\ref{fig:recombinant_qc}).
Evidence that this filter effectively removed artifacts can be seen
in the clear association between the AmpliSeq V1 kit and
recombination events that we identified as artifactual.

Different types of recombinants successfully removed are depicted
by the illustrative copying patterns in Figure~\ref{fig:copying_pattern_examples}:
(1) Q1, 686;
(2) Q3, MNS at 13423-13424 \cite{DeMaio2024};
(3) Q4, 22194 (possibly MAFFT alignment error \cite{Hunt2024}).
